---
title: "Checking Balance Improvement"
description: |
  Comparing hours with entering cruise traffic to hours without. Adjusting for calendar and weather indicators.
author:
  - name: Marie-Abèle Bind 
    url: https://scholar.harvard.edu/marie-abele
    affiliation: Biostatistics Center, Massachusetts General Hospital
    affiliation_url: https://biostatistics.massgeneral.org/faculty/marie-abele-bind-phd/
  - name: Marion Leroutier 
    url: https://www.parisschoolofeconomics.eu/en/leroutier-marion/work-in-progress/
    affiliation: Misum, Stockholm School of Economics
    affiliation_url: https://www.hhs.se/en/persons/l/leroutier-marion/
  - name: Léo Zabrocki 
    url: https://lzabrocki.github.io/
    affiliation: RFF-CMCC EIEE
    affiliation_url: https://www.eiee.org/
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
      toc: true
      toc_depth: 3
editor_options: 
  chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
# code chunk option
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  dev = "CairoPNG",
  dpi = 300
)
```

We adjust for calendar calendar indicator and weather confounding factors.

**Should you have any questions, need help to reproduce the analysis or find coding errors, please do not hesitate to contact us at leo.zabrocki@gmail.com and marion.leroutier@hhs.se.**

# Required Packages

We load the following packages:

```{r, message = FALSE, warning = FALSE}
# load required packages
library(knitr) # for creating the R Markdown document
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualization
library(dtplyr) # to speed up to dplyr
library(randChecks) # for randomization check
library(ggridges) # for ridge density plots
library(Cairo) # for printing custom police of graphs
library(patchwork) # combining plots
library(kableExtra) # for table formatting
```

We also load our custom **ggplot2** theme for graphs:

```{r}
# load ggplot custom theme
source(here::here(
  "inputs",
  "2.functions",
  "script_theme_tufte.R"
))
# define nice colors
my_blue <- "#0081a7"
my_orange <- "#fb8500"
```

# Preparing the Data

We load the matching and matched data and bind them together:

```{r}
# load Initial Data
data_matching <-
  readRDS(
    here::here(
      "inputs",
      "1.data",
      "1.hourly_data",
      "2.data_for_analysis",
      "1.matched_data",
      "matching_data.rds"
    )
  ) %>%
  mutate(dataset = "Initial Data")

# load matched data
data_matched <-
  readRDS(
    here::here(
      "inputs",
      "1.data",
      "1.hourly_data",
      "2.data_for_analysis",
      "1.matched_data",
      "matched_data_entry_cruise.rds"
    )
  ) %>%
  mutate(dataset = "Matched Data")

# bind the two datasets
data <- bind_rows(data_matched, data_matching) %>% dplyr::as_data_frame()
```

We also load the initial data to select the lags of variables at the **daily** level:

```{r}
# load initial data
data_daily <-
  readRDS(
    here::here(
      "inputs",
      "1.data",
      "1.hourly_data",
      "2.data_for_analysis",
      "0.main_data",
      "data_for_analysis_hourly.RDS"
    )
  ) %>%
  select(
    date,
    daily_mean_no2_sl_lag_1:daily_mean_o3_l_lag_2,
    daily_rainfall_height_dummy_lag_1:daily_humidity_average_lag_2
  ) %>% dplyr::as_data_frame()
```

We merge the `matched_data` with the `data_daily` :

```{r}
data <- data %>%
  left_join(., data_daily, by = "date")
```

We change labels of the `is_treated` variable :

```{r}
data <- data %>%
  mutate(is_treated = ifelse(is_treated == "TRUE", "True", "False"))
```

# Love Plots

### Continuous Weather Covariates

```{r fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
# compute figures for the love plot
data_weather_continuous <- data %>%
  select(
    dataset,
    is_treated,
    pair_number,
    contains("temperature"),
    contains("humidity"),
    contains("wind_speed")
  ) %>%
  pivot_longer(
    cols = -c(pair_number, is_treated, dataset),
    names_to = "variable",
    values_to = "values"
  ) %>%
  mutate(
    weather_variable = NA %>%
      ifelse(
        str_detect(variable, "temperature_average"),
        "Average Temperature",
        .
      ) %>%
      ifelse(
        str_detect(variable, "humidity_average"),
        "Humidity Average",
        .
      ) %>%
      ifelse(str_detect(variable, "wind_speed"), "Wind Speed", .)
  ) %>%
  mutate(
    time = 0 %>%
      ifelse(str_detect(variable, "lag_1"),-1, .) %>%
      ifelse(str_detect(variable, "lag_2"),-2, .) %>%
      ifelse(str_detect(variable, "lag_3"),-3, .) %>%
      ifelse(str_detect(variable, "lead_1"), 1, .) %>%
      ifelse(str_detect(variable, "lead_2"), 2, .) %>%
      ifelse(str_detect(variable, "lead_3"), 3, .)
  ) %>%
  filter(time<= 0) %>%
  mutate(level = "Hourly" %>%
           ifelse(str_detect(variable, "daily"), "Daily", .)) %>%
  filter(!is.na(time)) %>%
  mutate(level_time = paste(level, "Lag", time, sep = " ")) %>%
  select(dataset,
         pair_number,
         is_treated,
         weather_variable,
         level_time,
         values)

data_abs_difference_continuous_weather <-
  data_weather_continuous %>%
  group_by(dataset, weather_variable, level_time, is_treated) %>%
  summarise(mean_values = mean(values, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_values[2] - mean_values[1]))

data_sd_weather_continuous <-  data_weather_continuous %>%
  filter(is_treated == "True" & dataset == "Initial Data") %>%
  group_by( weather_variable, level_time, is_treated) %>%
  summarise(sd_treatment = sd(values, na.rm = TRUE)) %>%
  ungroup() %>%
  select(weather_variable, level_time, sd_treatment)

data_love_continuous_weather <-
  left_join(
    data_abs_difference_continuous_weather,
    data_sd_weather_continuous,
    by = c("weather_variable", "level_time")
  ) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment)) %>%
  mutate(level_time = fct_relevel(level_time, "Daily Lag -1", "Daily Lag -2", "Hourly Lag 0", "Hourly Lag -1", "Hourly Lag -2", "Hourly Lag -3"))



# make the graph
graph_love_plot_continuous_weather <-
  ggplot(
    data_love_continuous_weather,
    aes(
      y = fct_rev(level_time),
      x = standardized_difference,
      colour = fct_rev(dataset),
      shape = fct_rev(dataset),
    )
  ) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(
    size = 2,
    alpha = 0.8
  ) +
  scale_colour_manual(name = "Dataset: ", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset: ", values = c(16, 17)) +
  facet_wrap( ~ weather_variable, scales = "free_y") +
  xlab("Standardized Mean Differences") +
  ylab("") +
  theme_tufte()

# plot the graph
graph_love_plot_continuous_weather

# save the graph
ggsave(
  graph_love_plot_continuous_weather,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_continuous_weather.pdf"
  ),
  width = 55,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

### Categorical Weather Covariates

```{r fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
# compute figures for the love plot
data_weather_categorical <- data %>%
  select(
    dataset,
    is_treated,
    contains("rainfall_height_dummy"),
    contains("wind_direction_east_west")
  ) %>%
  mutate_at(vars(contains("rainfall")), ~ ifelse(. == 1, "True", "False")) %>%
  mutate_all( ~ as.character(.)) %>%
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "variable",
    values_to = "values"
  ) %>%
  # group by is_treated, variable and values
  group_by(dataset, is_treated, variable, values) %>%
  # compute the number of observations
  summarise(n = n()) %>%
  # compute the proportion
  mutate(freq = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(
    weather_variable = NA %>%
      ifelse(str_detect(variable, "wind"), "Wind Direction", .) %>%
      ifelse(str_detect(variable, "rainfall"), "Rainfall Dummy", .)
  ) %>%
  mutate(
    time = 0 %>%
      ifelse(str_detect(variable, "lag_1"),-1, .) %>%
      ifelse(str_detect(variable, "lag_2"),-2, .) %>%
      ifelse(str_detect(variable, "lag_3"),-3, .) %>%
      ifelse(str_detect(variable, "lead_1"), 1, .) %>%
      ifelse(str_detect(variable, "lead_2"), 2, .) %>%
      ifelse(str_detect(variable, "lead_3"), 3, .)
  ) %>%
  filter(time<= 0) %>%
  mutate(level = "Hourly" %>%
           ifelse(str_detect(variable, "daily"), "Daily", .)) %>%
  filter(!is.na(time)) %>%
  mutate(variable = paste(level, weather_variable, "\nLag", time, sep = " ")) %>%
  select(dataset, is_treated, weather_variable, variable, values, freq) %>%
  pivot_wider(names_from = is_treated, values_from = freq) %>%
  mutate(abs_difference = abs(`True` - `False`)) %>%
  filter(values != "False" & values != "West") 

# create the figure for wind direction
graph_love_plot_wind_direction <- data_weather_categorical %>%
  filter(weather_variable == "Wind Direction") %>%
  mutate(variable = fct_relevel(variable, "Daily Wind Direction \nLag -1", "Daily Wind Direction \nLag -2", "Hourly Wind Direction \nLag 0", "Hourly Wind Direction \nLag -1", "Hourly Wind Direction \nLag -2", "Hourly Wind Direction \nLag -3")) %>%
  ggplot(., aes(
    y = fct_rev(values),
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(
    size = 2,
    alpha = 0.8
  ) +
  scale_colour_manual(name = "Dataset: ", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset: ", values = c(16, 17)) +
  facet_wrap( ~ variable, scales = "free_y", ncol = 4) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte()

# print the figure for wind direction
graph_love_plot_wind_direction

# save the figure for wind direction
ggsave(
  graph_love_plot_wind_direction,
  filename = here::here(
      "inputs",
      "3.outputs",
      "1.hourly_analysis",
      "2.experiment_cruise",
      "1.checking_matching_procedure",
    "graph_love_plot_wind_direction.pdf"
  ),
  width = 60,
  height = 30,
  units = "cm",
  device = cairo_pdf
)

# create the figure for rainfall dummy
graph_love_plot_rainfall <- data_weather_categorical %>%
  filter(weather_variable == "Rainfall Dummy") %>%
  mutate(variable = fct_relevel(variable, "Daily Rainfall Dummy \nLag -1", "Daily Rainfall Dummy \nLag -2", "Hourly Rainfall Dummy \nLag 0", "Hourly Rainfall Dummy \nLag -1", "Hourly Rainfall Dummy \nLag -2", "Hourly Rainfall Dummy \nLag -3")) %>%
  ggplot(., aes(
    y = fct_rev(variable),
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(
    size = 2,
    alpha = 0.8
  ) +
  scale_colour_manual(name = "Dataset: ", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset: ", values = c(16, 17)) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte()

# print the figure for rainfall dummy
graph_love_plot_rainfall

# save the figure for rainfall dummy
ggsave(
  graph_love_plot_rainfall,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_rainfall.pdf"
  ),
  width = 30,
  height = 30,
  units = "cm",
  device = cairo_pdf
)
```

### Pollutants

```{r fig.width=10, fig.height=10, layout="l-page", code_folding="Please show me the code!"}
# compute figures for the love plot
data_pollutants <- data %>%
  select(
    dataset,
    is_treated,
    pair_number,
    contains("no2"),
    contains("o3"),
    contains("pm10"),
    contains("pm25"),
    contains("so2")
  ) %>%
  pivot_longer(
    cols = -c(pair_number, is_treated, dataset),
    names_to = "variable",
    values_to = "values"
  ) %>%
  mutate(
    pollutant = NA %>%
      ifelse(str_detect(variable, "no2_l"), "NO2 Longchamp", .) %>%
      ifelse(str_detect(variable, "no2_sl"), "NO2 Saint-Louis", .) %>%
      ifelse(str_detect(variable, "o3"), "O3 Lonchamp", .) %>%
      ifelse(str_detect(variable, "pm10_l"), "PM10 Longchamp", .) %>%
      ifelse(str_detect(variable, "pm10_sl"), "PM10 Saint-Louis", .) %>%
      ifelse(str_detect(variable, "pm25"), "PM2.5 Longchamp", .) %>%
      ifelse(str_detect(variable, "so2"), "SO2 Longchamp", .)
  ) %>%
  mutate(
    time = 0 %>%
      ifelse(str_detect(variable, "lag_1"),-1, .) %>%
      ifelse(str_detect(variable, "lag_2"),-2, .) %>%
      ifelse(str_detect(variable, "lag_3"),-3, .) %>%
      ifelse(str_detect(variable, "lead_1"), 1, .) %>%
      ifelse(str_detect(variable, "lead_2"), 2, .) %>%
      ifelse(str_detect(variable, "lead_3"), 3, .)
  ) %>%
  filter(time<= 0) %>%
  mutate(level = "Hourly" %>%
           ifelse(str_detect(variable, "daily"), "Daily", .)) %>%
  filter(!is.na(time)) %>%
  mutate(level_time = paste(level, "Lag", time, sep = " ")) %>%
  select(dataset,
         pair_number,
         is_treated,
         pollutant,
         level_time,
         values)

data_abs_difference_pollutants <- data_pollutants %>%
  group_by(dataset, pollutant, level_time, is_treated) %>%
  summarise(mean_values = mean(values, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_values[2] - mean_values[1]))

data_sd_pollutants <-  data_pollutants %>%
  filter(is_treated == "True" & dataset == "Initial Data") %>%
  group_by(pollutant, level_time, is_treated) %>%
  summarise(sd_treatment = sd(values, na.rm = TRUE)) %>%
  ungroup() %>%
  select(pollutant, level_time, sd_treatment)

data_love_pollutants <-
  left_join(data_abs_difference_pollutants,
            data_sd_pollutants,
            by = c("pollutant", "level_time")) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment)) %>%
  mutate(level_time = fct_relevel(level_time, "Daily Lag -1", "Daily Lag -2", "Hourly Lag 0", "Hourly Lag -1", "Hourly Lag -2", "Hourly Lag -3"))

# create the graph
graph_love_plot_pollutants <-
  ggplot(data_love_pollutants,
         aes(
           y = fct_rev(level_time),
           x = standardized_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(
    size = 2,
    alpha = 0.8
  ) +
  scale_colour_manual(name = "Dataset: ", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset: ", values = c(16, 17)) +
  facet_wrap( ~ pollutant, scales = "free_y") +
  xlab("Standardized Mean Differences") +
  ylab("") +
  theme_tufte()

# print the graph
graph_love_plot_pollutants

# save the graph
ggsave(
  graph_love_plot_pollutants,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_pollutants.pdf"
  ),
  width = 60,
  height = 40,
  units = "cm",
  device = cairo_pdf
)
```


### Vessels Traffic

```{r fig.width=10, fig.height=10, layout="l-page", code_folding="Please show me the code!"}
# compute figures for the love plot
data_tonnage <-  data %>%
  # select relevant variables
  select(
    dataset,
    pair_number,
    is_treated,
    contains("total_gross_tonnage_entry_cruise"),
    contains("total_gross_tonnage_exit_cruise"),
    contains("total_gross_tonnage_entry_ferry"),
    contains("total_gross_tonnage_exit_ferry"),
    contains("total_gross_tonnage_entry_other_vessels"),
    contains("total_gross_tonnage_exit_other_vessels")
  ) %>%
  # transform data in long format
  pivot_longer(
    cols = -c(dataset, pair_number, is_treated),
    names_to = "variable",
    values_to = "tonnage"
  ) %>%
  # create vessel type variable
  mutate(
    vessel_type = NA %>%
      ifelse(str_detect(variable, "cruise"), "Cruise", .) %>%
      ifelse(str_detect(variable, "ferry"), "Ferry", .) %>%
      ifelse(str_detect(variable, "other_vessels"), "Other Vessels", .)
  ) %>%
  # create the day variable
  mutate(entry_exit = NA %>%
           ifelse(str_detect(variable, "entry"), "Entry", .) %>%
           ifelse(str_detect(variable, "exit"), "Exit", .)) %>%
  mutate(
    time = 0 %>%
      ifelse(str_detect(variable, "lag_1"),-1, .) %>%
      ifelse(str_detect(variable, "lag_2"),-2, .) %>%
      ifelse(str_detect(variable, "lag_3"),-3, .) %>%
      ifelse(str_detect(variable, "lead_1"), 1, .) %>%
      ifelse(str_detect(variable, "lead_2"), 2, .) %>%
      ifelse(str_detect(variable, "lead_3"), 3, .)
  ) %>%
  select(dataset,
         pair_number,
         vessel_type,
         is_treated,
         entry_exit,
         time,
         tonnage)

data_abs_difference_tonnage <- data_tonnage %>%
  group_by(dataset, vessel_type, entry_exit, time, is_treated) %>%
  summarise(mean_tonnage = mean(tonnage, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_tonnage[2] - mean_tonnage[1]))

data_sd_tonnage <-  data_tonnage %>%
  filter(is_treated == "True" & dataset == "Initial Data") %>%
  group_by(vessel_type, entry_exit, time, is_treated) %>%
  summarise(sd_treatment = sd(tonnage, na.rm = TRUE)) %>%
  ungroup() %>%
  select(vessel_type, entry_exit, time, sd_treatment)

data_love_tonnage <-
  left_join(
    data_abs_difference_tonnage,
    data_sd_tonnage,
    by = c("vessel_type", "entry_exit", "time")
  ) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment)) %>%
  filter(!(vessel_type == "Cruise" & time == 0))

# create the graph
graph_love_plot_tonnage <-
  ggplot(data_love_tonnage,
         aes(
           x = standardized_difference,
           y = as.factor(time),
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(
    size = 2,
    alpha = 0.8
  ) +
  scale_colour_manual(name = "Dataset: ", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset: ", values = c(16, 17)) +
  facet_grid(entry_exit ~ vessel_type) +
  xlab("Standardized Mean Differences") +
  ylab("Hour") +
  theme_tufte()

# print the graph
graph_love_plot_tonnage

# save the graph
ggsave(
  graph_love_plot_tonnage,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_tonnage.pdf"
  ),
  width = 40,
  height = 40,
  units = "cm",
  device = cairo_pdf
)
```

### Road Traffic

```{r fig.width=10, fig.height=10, layout="l-page", code_folding="Please show me the code!"}
# compute figures for the love plot
data_road <- data %>%
  # select relevant variables
  select(dataset,
         pair_number,
         is_treated,
         contains("road_traffic_flow")) %>%
  # transform data in long format
  pivot_longer(
    cols = -c(dataset, pair_number, is_treated),
    names_to = "road_traffic",
    values_to = "value"
  ) %>%
  mutate(
    time = 0 %>%
      ifelse(str_detect(road_traffic, "lag_1"),-1, .) %>%
      ifelse(str_detect(road_traffic, "lag_2"),-2, .) %>%
      ifelse(str_detect(road_traffic, "lag_3"),-3, .) %>%
      ifelse(str_detect(road_traffic, "lead_1"), 1, .) %>%
      ifelse(str_detect(road_traffic, "lead_2"), 2, .) %>%
      ifelse(str_detect(road_traffic, "lead_3"), 3, .)
  ) %>%
  select(dataset, pair_number, road_traffic, is_treated, time, value)

data_abs_difference_road <- data_road %>%
  group_by(dataset, road_traffic, time, is_treated) %>%
  summarise(mean_value = mean(value, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_value[2] - mean_value[1]))

data_sd_tonnage <-  data_road %>%
  filter(is_treated == "True" & dataset == "Initial Data") %>%
  group_by(road_traffic, time, is_treated) %>%
  summarise(sd_treatment = sd(value, na.rm = TRUE)) %>%
  ungroup() %>%
  select(road_traffic, time, sd_treatment)

data_love_road <-
  left_join(data_abs_difference_road,
            data_sd_tonnage,
            by = c("road_traffic", "time")) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment))


# create the graph
graph_love_plot_road <-
  ggplot(
    data_love_road,
    aes(
      x = standardized_difference,
      y = as.factor(time),
      colour = fct_rev(dataset),
      shape = fct_rev(dataset)
    )
  ) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(16, 17)) +
  xlab("Standardized Mean Differences") +
  ylab("Hour") +
  theme_tufte() 

# print the graph
graph_love_plot_road


# save the graph
ggsave(
  graph_love_plot_road,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_road.pdf"
  ),
  width = 25,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

### Calendar Indicators

Create the relevant data:

```{r fig.width=10, fig.height=10, layout="l-page", code_folding="Please show me the code!"}
# compute figures for the love plot
data_calendar <- data %>%
  mutate(weekday = lubridate::wday(date, abbr = FALSE, label = TRUE)) %>%
  select(dataset,
         is_treated,
         hour,
         weekday,
         holidays_dummy,
         bank_day_dummy,
         month,
         year) %>%
  mutate_at(vars(holidays_dummy, bank_day_dummy),
            ~ ifelse(. == 1, "True", "False")) %>%
  mutate_all( ~ as.character(.)) %>%
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "variable",
    values_to = "values"
  ) %>%
  # group by is_treated, variable and values
  group_by(dataset, is_treated, variable, values) %>%
  # compute the number of observations
  summarise(n = n()) %>%
  # compute the proportion
  mutate(freq = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(
    calendar_variable = NA %>%
      ifelse(str_detect(variable, "hour"), "Hour", .) %>%
      ifelse(str_detect(variable, "weekday"), "Day of the Week", .) %>%
      ifelse(str_detect(variable, "holidays_dummy"), "Holidays", .) %>%
      ifelse(str_detect(variable, "bank_day_dummy"), "Bank Day", .) %>%
      ifelse(str_detect(variable, "month"), "Month", .) %>%
      ifelse(str_detect(variable, "year"), "Year", .)
  ) %>%
  select(dataset, is_treated, calendar_variable, values, freq) %>%
  pivot_wider(names_from = is_treated, values_from = freq) %>%
  mutate(abs_difference = abs(`True` - `False`)) %>%
  filter(values != "False")
```

Plot for hours:

```{r fig.width=10, fig.height=6, layout="l-page", code_folding="Please show me the code!"}
# graph for hour
graph_love_plot_hour <- data_calendar %>%
  filter(calendar_variable == "Hour") %>%
  ggplot(., aes(
    y = as.factor(as.numeric(values)),
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(16, 17)) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte()

# print the plot
graph_love_plot_hour

# save the plot
ggsave(
  graph_love_plot_hour,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_hour.pdf"
  ),
  width = 30,
  height = 30,
  units = "cm",
  device = cairo_pdf
)
```

Plot for bank days and holidays:

```{r fig.width=4, fig.height=4, layout="l-page", code_folding="Please show me the code!"}
# graph for bank days and holidays
graph_love_plot_bank_holidays <- data_calendar %>%
  filter(calendar_variable %in% c("Bank Day", "Holidays")) %>%
  ggplot(., aes(
    y = values,
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(16, 17)) +
  facet_wrap( ~ calendar_variable) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() 

# print the plot
graph_love_plot_bank_holidays

# save the plot
ggsave(
  graph_love_plot_bank_holidays,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_bank_holidays.pdf"
  ),
  width = 30,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

Plot for days of the week:

```{r fig.width=10, fig.height=6, layout="l-page", code_folding="Please show me the code!"}
# graph for weekdays
graph_love_plot_weekday <- data_calendar %>%
  filter(calendar_variable == "Day of the Week") %>%
  mutate(
    values = fct_relevel(
      values,
      "Monday",
      "Tuesday",
      "Wednesday",
      "Thursday",
      "Friday",
      "Saturday",
      "Sunday"
    )
  ) %>%
  ggplot(., aes(
    y = fct_rev(values),
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(16, 17)) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() 

# print the plot
graph_love_plot_weekday

# save the plot
ggsave(
  graph_love_plot_weekday,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_weekday.pdf"
  ),
  width = 30,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

Plot for months:

```{r fig.width=10, fig.height=6, layout="l-page", code_folding="Please show me the code!"}
# graph for month
graph_love_plot_month <- data_calendar %>%
  filter(calendar_variable == "Month") %>%
  mutate(
    values = fct_relevel(
      values,
      "January",
      "February",
      "March",
      "April",
      "May",
      "June",
      "July",
      "August",
      "September",
      "October",
      "November",
      "December"
    )
  ) %>%
  ggplot(., aes(
    y = fct_rev(values),
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(16, 17)) +
  ggtitle("Month") +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() 

# print the plot
graph_love_plot_month

# save the plot
ggsave(
  graph_love_plot_month,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_month.pdf"
  ),
  width = 30,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

Plot for years:

```{r fig.width=10, fig.height=6, layout="l-page", code_folding="Please show me the code!"}
# graph for year
graph_love_plot_year <- data_calendar %>%
  filter(calendar_variable == "Year") %>%
  ggplot(., aes(
    y = as.factor(as.numeric(values)),
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(16, 17)) +
  ggtitle("Year") +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte()

# print the graph
graph_love_plot_year

# save the plot
ggsave(
  graph_love_plot_year,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_love_plot_year.pdf"
  ),
  width = 30,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

# Overall Balance Improvement

We finally plot the distribution of standardized mean differences for continuous covariates or the absolute percentage points differences for categorical covariates between treated and control units before and after matching.

### Continuous Covariates

```{r fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
# we select the dataset indicator and the standardized difference
data_love_tonnage <- data_love_tonnage %>%
  ungroup() %>%
  filter(time < 0) %>%
  select(dataset, standardized_difference)

data_love_pollutants <- data_love_pollutants %>%
  ungroup() %>%
  filter(parse_number(as.character(level_time))  < 0) %>%
  select(dataset, standardized_difference)

data_love_continuous_weather <- data_love_continuous_weather %>%
  ungroup() %>%
  filter(parse_number(as.character(level_time))  <= 0) %>%
  select(dataset, standardized_difference)

data_love_road <- data_love_road %>%
  ungroup() %>%
  filter(time <= 0) %>%
  select(dataset, standardized_difference)

data_continuous_love <-
  bind_rows(data_love_tonnage, data_love_pollutants) %>%
  bind_rows(., data_love_continuous_weather) %>%
  bind_rows(., data_love_road)

# create the graph
graph_boxplot_continuous_balance_improvement <-
  ggplot(data_continuous_love,
         aes(x = dataset, y = standardized_difference)) +
  geom_boxplot(colour = my_blue) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  xlab("Dataset") +
  ylab("Standardized \nMean Differences") +
  theme_tufte()

# print the graph
graph_boxplot_continuous_balance_improvement

# save the graph
ggsave(
  graph_boxplot_continuous_balance_improvement,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_boxplot_continuous_balance_improvement.pdf"
  ),
  width = 30,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

### Categorical Covariates

```{r fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
# we select the dataset indicator and the standardized difference
data_calendar <- data_calendar %>%
  ungroup() %>%
  # we remove the month variable
  #filter(calendar_variable != "Month") %>%
  select(dataset, abs_difference) %>%
  drop_na()

data_weather_categorical <- data_weather_categorical %>%
  ungroup() %>%
  select(dataset, abs_difference)

data_categorical_love <-
  bind_rows(data_calendar, data_weather_categorical) %>%
  drop_na()

# create the graph
graph_boxplot_categorical_balance_improvement <-
  ggplot(data_categorical_love, aes(x = dataset, y = abs_difference)) +
  geom_boxplot(colour = my_blue) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  xlab("Dataset") +
  ylab("Absolute Difference \nin Percentage Points") +
  theme_tufte()

# print the graph
graph_boxplot_categorical_balance_improvement

# save the graph
ggsave(
  graph_boxplot_categorical_balance_improvement,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_boxplot_categorical_balance_improvement.pdf"
  ),
  width = 30,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

We combine the two previous plots in a single figure:

```{r, code_folding="Please show me the code!"}
# combine plots
graph_overall_balance <-
  graph_boxplot_continuous_balance_improvement + graph_boxplot_categorical_balance_improvement +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 30, face = "bold"))

# save the plot
ggsave(
  graph_overall_balance,
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "graph_overall_balance_hourly.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

And we compute the overall figures for imbalance before and after matching:

```{r, code_folding="Please show me the code!"}
# compute average imbalance before and after matching
data_categorical_love <- data_categorical_love %>%
  mutate(Type = "Categorical (Difference in Percentage Points)") %>%
  rename(standardized_difference = abs_difference)
  
data_continuous_love %>%
  mutate(Type = "Continuous (Standardized Difference)") %>%
  bind_rows(data_categorical_love) %>%
  group_by(Type, dataset) %>%
  summarise("Mean Imbalance" = round(mean(standardized_difference), 2)) %>%
  rename(Dataset = dataset) %>%
  knitr::kable(., align = c("l", "l", "c")) %>%
  kable_styling(position = "center")
```


# Randomization Check for Covariate Balance

Finally, we implement the procedure of [Zach Branson (2021)](https://drive.google.com/file/d/16G59khS-HVUjsSYY3NTT3yOq-Y-26la4/view) to carry out a randomization check to test whether daily observations are as-if randomized according to observed covariates.

```{r, eval = FALSE}
# we select covariates
data_covs <- data_matched %>%
  select(
    pair_number,
    is_treated,
    total_gross_tonnage_ferry,
    total_gross_tonnage_entry_ferry,
    total_gross_tonnage_entry_other_vessels,
    total_gross_tonnage_exit_cruise,
    total_gross_tonnage_exit_ferry,
    total_gross_tonnage_exit_other_vessels,
    total_gross_tonnage_cruise_lag_1,
    total_gross_tonnage_ferry_lag_1,
    total_gross_tonnage_entry_ferry_lag_1,
    total_gross_tonnage_entry_other_vessels_lag_1,
    total_gross_tonnage_exit_cruise_lag_1,
    total_gross_tonnage_exit_ferry_lag_1,
    total_gross_tonnage_exit_other_vessels_lag_1,
    total_gross_tonnage_cruise_lag_2,
    total_gross_tonnage_ferry_lag_2,
    total_gross_tonnage_entry_ferry_lag_2,
    total_gross_tonnage_entry_other_vessels_lag_2,
    total_gross_tonnage_exit_cruise_lag_2,
    total_gross_tonnage_exit_ferry_lag_2,
    total_gross_tonnage_exit_other_vessels_lag_2,
    total_gross_tonnage_cruise_lag_3,
    total_gross_tonnage_ferry_lag_3,
    total_gross_tonnage_entry_ferry_lag_3,
    total_gross_tonnage_entry_other_vessels_lag_3,
    total_gross_tonnage_exit_cruise_lag_3,
    total_gross_tonnage_exit_ferry_lag_3,
    total_gross_tonnage_exit_other_vessels_lag_3,
    temperature_average:wind_speed,
    wind_direction_east_west,
    temperature_average_lag_1:wind_speed_lag_1,
    wind_direction_east_west_lag_1,
    temperature_average_lag_2:wind_speed_lag_2,
    wind_direction_east_west_lag_2,
    temperature_average_lag_3:wind_speed_lag_3,
    wind_direction_east_west_lag_3,
    hour,
    weekday,
    holidays_dummy,
    bank_day_dummy,
    month,
    year
  ) %>% drop_na()

# recode some variables
data_covs <- data_covs %>%
  mutate(is_treated = ifelse(is_treated == TRUE, 1, 0)) %>%
  mutate_at(
    vars(
      wind_direction_east_west ,
      wind_direction_east_west_lag_1,
      wind_direction_east_west_lag_2,
      wind_direction_east_west_lag_3
    ),
    ~ ifelse(. == "West", 1, 0)
  ) %>%
  fastDummies::dummy_cols(., select_columns = c("year", "month", "weekday", "hour")) %>%
  select(-c("year", "month", "weekday", "hour"))

# format data for asIfRandPlot() function
pair_number <- as.numeric(data_covs$pair_number)
is_treated <- data_covs$is_treated

data_covs <- data_covs %>%
  select(-pair_number, -is_treated) %>%
  select(
    total_gross_tonnage_ferry,
    total_gross_tonnage_entry_ferry,
    total_gross_tonnage_entry_other_vessels,
    total_gross_tonnage_exit_cruise,
    total_gross_tonnage_exit_other_vessels,
    total_gross_tonnage_cruise_lag_1,
    total_gross_tonnage_ferry_lag_1,
    total_gross_tonnage_exit_other_vessels_lag_1,
    total_gross_tonnage_cruise_lag_2,
    total_gross_tonnage_ferry_lag_2,
    total_gross_tonnage_exit_other_vessels_lag_1,
    total_gross_tonnage_cruise_lag_2,
    total_gross_tonnage_ferry_lag_2,
    total_gross_tonnage_entry_ferry_lag_2,
    total_gross_tonnage_entry_ferry_lag_2,
    total_gross_tonnage_entry_ferry_lag_2,
    total_gross_tonnage_entry_other_vessels_lag_2,
    total_gross_tonnage_cruise_lag_3,
    total_gross_tonnage_ferry_lag_3,
    total_gross_tonnage_entry_other_vessels_lag_3,
    total_gross_tonnage_exit_cruise_lag_3,
    total_gross_tonnage_exit_ferry_lag_3,
    total_gross_tonnage_exit_other_vessels_lag_3,
    temperature_average,
    humidity_average,
    wind_speed,
    wind_direction_east_west,
    temperature_average_lag_1:wind_direction_east_west_lag_1,
    temperature_average_lag_2,
    humidity_average_lag_2:wind_direction_east_west_lag_3,-year_2018,
    -month_December,
    -weekday_Sunday,
    -hour_0,
    -bank_day_dummy
  )

# run balance check
asIfRandPlot(
  X.matched = data_covs,
  indicator.matched = is_treated,
  assignment = c("complete", "blocked"),
  subclass = pair_number,
  perms = 1000
)
```

```{r, echo = FALSE, out.width = "800%", layout="l-page", fig.align = "center"}
knitr::include_graphics(
  here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiment_cruise",
    "1.checking_matching_procedure",
    "ri_balance_check.png"
  )
)
```







