---
title: "Exploraty Data Analysis of the Effect of Road Traffic on NO2 Concentration in Marseille"
description: |
  Comparing weekdays to weekend.
author:
  - name: Marie-Abèle Bind 
    url: https://scholar.harvard.edu/marie-abele
    affiliation: Biostatistics Center, Massachusetts General Hospital
    affiliation_url: https://biostatistics.massgeneral.org/faculty/marie-abele-bind-phd/
  - name: Marion Leroutier 
    url: https://www.parisschoolofeconomics.eu/en/leroutier-marion/work-in-progress/
    affiliation: Misum, Stockholm School of Economics
    affiliation_url: https://www.hhs.se/en/persons/l/leroutier-marion/
  - name: Léo Zabrocki 
    url: https://lzabrocki.github.io/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/fr/zabrocki-leo/
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
      toc: true
      toc_depth: 3
editor_options: 
  chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
# code chunk option
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  dev = "CairoPNG",
  dpi = 300
)
```

In this document

**Should you have any questions, need help to reproduce the analysis or find coding errors, please do not hesitate to contact us at leo.zabrocki@psemail.eu and marion.leroutier@hhs.se.**

# Required Packages

To reproduce exactly the `2_9_daily_regression_analysis.html` document, we first need to have installed:

* the [R](https://www.r-project.org/) programming language on your computer 
* [RStudio](https://rstudio.com/), an integrated development environment for R, which will allow you to knit the `2_9_daily_regression_analysis.Rmd` file and interact with the R code chunks
* the [R Markdown](https://rmarkdown.rstudio.com/) package
* and the [Distill](https://rstudio.github.io/distill/) package which provides the template of this document. 

Once everything is set up, we have to load the following packages:

```{r}
# load required packages
library(knitr) # for creating the R Markdown document
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualization
library(broom) # for cleaning regression outputs
library(randChecks) # for randomization check
library(lmtest) # for modifying regression standard errors
library(sandwich) # for robust and cluster robust standard errors
library(kableExtra) # for table formatting
library(Cairo) # for printing custom police of graphs
```

We load our custom `ggplot2` theme for graphs:

```{r}
# load ggplot custom theme
source(here::here("inputs",
                  "2.functions",
                  "script_theme_tufte.R"))
# define nice colors
my_blue <- "#0081a7"
my_orange <- "#fb8500"
```


Finally, we ran a simple time-stratified regression model on the matching data to see how the results differ with those found in our matched data analysis. We only adjust for calendar indicator and weather covariates measured at time t. Again, we ran this procedure on an Amazon Web Services virtual computer (EC2 t3.2xlarge). The code we used is in the script `script_aws_regression_matching_data.R` located in the **1.experiment_entry_cruise/script_aws** folder. It takes several minutes to run.

We load the outputs of the regression model and plot the 95% confidence intervals:

```{r, layout="l-body-outset", fig.width=30, fig.height=12, code_folding="Please show me the code!"}
# load regression outputs
data_reg_analysis <- readRDS(
  here::here(
    "inputs",
    "1.data",
    "1.hourly_data",
    "2.data_for_analysis",
    "1.matched_data",
    "1.experiments_cruise",
    "1.experiment_entry_cruise",
    "data_matching_regression.rds"
  )
) %>%
  # compute 95% confidence intervals using the standard error
  mutate(ci_lower_95 = estimate - 1.96 * std.error,
         ci_upper_95 = estimate + 1.96 * std.error) %>%
  select(-std.error) %>%
  mutate(analysis = "Outcome Regression") %>%
  rename(mean_difference = estimate)

data_neyman <- readRDS(
  here::here(
    "inputs",
    "1.data",
    "1.hourly_data",
    "2.data_for_analysis",
    "1.matched_data",
    "1.experiments_cruise",
    "1.experiment_entry_cruise",
    "data_neyman.rds"
  )
)

data_neyman <- data_neyman %>%
  mutate(analysis = "Constrained Pair Matching")

data_neyman_reg <- bind_rows(data_neyman, data_reg_analysis)

# make the graph

# create an indicator to alternate shading of confidence intervals
data_neyman_reg <- data_neyman_reg %>%
  arrange(pollutant, time) %>%
  mutate(stripe = ifelse((time %% 2) == 0, "Grey", "White")) %>%
  ungroup()

# make the graph
graph_reg_analysis <-
  ggplot(
    data_neyman_reg,
    aes(
      x = as.factor(time),
      y = mean_difference,
      ymin = ci_lower_95,
      ymax = ci_upper_95,
      colour = analysis,
      shape = analysis
    )
  ) +
  geom_rect(
    aes(fill = stripe),
    xmin = as.numeric(as.factor(data_neyman_reg$time)) - 0.42,
    xmax = as.numeric(as.factor(data_neyman_reg$time)) + 0.42,
    ymin = -Inf,
    ymax = Inf,
    color = NA,
    alpha = 0.4
  ) +
  geom_hline(yintercept = 0, color = "black") +
  geom_pointrange(position = position_dodge(width = 1), size = 1.2) +
  scale_shape_manual(name = "Analysis:", values = c(16, 17)) +
  scale_color_manual(name = "Analysis:", values = c(my_orange, my_blue)) +
  facet_wrap( ~ pollutant, ncol = 4) +
  scale_fill_manual(values = c('grey90', "white")) +
  guides(fill = FALSE) +
  ylab("Average Difference \nin Concentrations (µg/m³)") + xlab("Day") +
  theme_tufte()

# display the graph
graph_reg_analysis

# save the graph
ggsave(
  graph_reg_analysis + labs(title = NULL),
  filename = here::here(
    "inputs",
    "3.outputs",
    "1.hourly_analysis",
    "2.experiments_cruise",
    "1.experiment_entry_cruise",
    "2.matching_results",
    "graph_matching_regression.pdf"
  ),
  width = 30,
  height = 15,
  units = "cm",
  device = cairo_pdf
)
```

And we display below the point estimates with the 95% confidence intervals:

```{r, code_folding="Please show me the code!"}
data_reg_analysis %>%
  select(pollutant, time, mean_difference, ci_lower_95, ci_upper_95) %>%
  mutate_at(vars(mean_difference:ci_upper_95), ~ round(., 1)) %>%
  rename(
    "Pollutant" = pollutant,
    "Time" = time,
    "Point Estimate" = mean_difference,
    "Lower Bound of the 95% Confidence Interval" = ci_lower_95,
    "Upper Bound of the 95% Confidence Interval" = ci_upper_95
  ) %>%
  rmarkdown::paged_table(.)
```