---
title: "Exploraty Data Analysis of the Effect of Road Traffic on NO2 Concentration in Marseille"
description: |
  Comparing weekdays to weekend.
author:
  - name: Marie-Abèle Bind 
    url: https://scholar.harvard.edu/marie-abele
    affiliation: Biostatistics Center, Massachusetts General Hospital
    affiliation_url: https://biostatistics.massgeneral.org/faculty/marie-abele-bind-phd/
  - name: Marion Leroutier 
    url: https://www.parisschoolofeconomics.eu/en/leroutier-marion/work-in-progress/
    affiliation: Misum, Stockholm School of Economics
    affiliation_url: https://www.hhs.se/en/persons/l/leroutier-marion/
  - name: Léo Zabrocki 
    url: https://lzabrocki.github.io/
    affiliation: RFF-CMCC EIEE
    affiliation_url: https://www.eiee.org/
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
      toc: true
      toc_depth: 3
editor_options: 
  chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
# code chunk option
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  dev = "CairoPNG",
  dpi = 300
)
```

In this document, we carry out an exploratory data analysis to investigate the parallel evolution of road traffic and nitrogen dioxide (NO$_{2}$) concentrations in Marseille. We focus on NO$_{2}$ as it is locally emitted by cars. We carry out a simple analysis where compare the concentration of NO$_{2}$ between weekdays and weekends as road traffic tends to decrease on Saturday and Sunday. The weekend/weedays contrast can be seen as a form of arbitrary variation in road traffic. We do not check the effects on other pollutants because they are less emitted locally.

In the following sections, we proceed as follows:

* We show that road traffic is indeed lower on weekends and that NO$_{2}$ concentrations fall at the end of the week.
* We check that weather covariates are balanced across weekdays and weekends.
* We compute the difference in NO$_{2}$ concentrations between weekdays and weekends.

**Should you have any questions, need help to reproduce the analysis or find coding errors, please do not hesitate to contact us at leo.zabrocki@gmail.com and marion.leroutier@hhs.se.**

# Required Packages

We load the following packages:

```{r}
# load required packages
library(knitr) # for creating the R Markdown document
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualization
library(kableExtra) # for table formatting
library(Cairo) # for printing custom police of graphs
```

We load our custom `ggplot2` theme for graphs:

```{r}
# load ggplot custom theme
source(here::here("inputs",
                  "2.functions",
                  "script_theme_tufte.R"))
# define nice colors
my_blue <- "#0081a7"
my_orange <- "#fb8500"
```

# Data Loading and Formatting

First, we load the daily data:

```{r}
# load data
data <-
  readRDS(
    here(
      "inputs",
      "1.data",
      "2.daily_data",
      "2.data_for_analysis",
      "0.main_data",
      "data_for_analysis_daily.RDS"
    )
  )
```

We select relevant variables for our analysis:

```{r}
data <- data %>%
  select(
    # date variable
    "date",
    # pollutants
    "mean_no2_l",
    "mean_no2_sl",
    # maritime traffic variables
    "total_gross_tonnage_cruise",
    # weather parameters
    "temperature_average",
    "rainfall_height_dummy",
    "humidity_average",
    "wind_speed",
    "wind_direction_categories",
    # road traffic variables
    "road_traffic_flow_all",
    "road_occupancy_rate",
    # calendar indicators
    "weekday",
    "weekend",
    "holidays_dummy",
    "bank_day_dummy",
    "month",
    "year"
  )
```

# Exploratory Data Analysis

### Road Traffic Variation by Day of the Week

We explore the patterns of road traffic. It is important to keep in mind that:

* We only use six stations in Marseille to create aggregated measures of traffic.
* Some stations have an important number of missing values.
* Data are available only from 2011-01-01	to 2016-10-02.

We first plot the average road traffic flow and occupancy rate by day of the week :

```{r, fig.width=10, fig.height=4, layout="l-page", code_folding="Please show me the code!"}
# make the graph
graph_road_traffic_wd <- data %>%
  select(weekday, road_traffic_flow_all, road_occupancy_rate) %>%
  pivot_longer(cols = -c(weekday), names_to = "traffic_variable", values_to = "value") %>%
  mutate(traffic_variable = ifelse(traffic_variable == "road_traffic_flow_all", "Flow of Vehicles (N)", "Occupancy Rate (%)")) %>%
  group_by(weekday, traffic_variable) %>%
  summarise(mean_value = mean(value, na.rm = TRUE)) %>%
  ggplot(., aes(x = weekday, y = mean_value)) +
  geom_segment(aes(x = weekday, xend = weekday, y = 0, yend = mean_value)) +
  geom_point(shape = 21, size = 4, colour = "black", fill = my_blue) +
  facet_wrap(~ traffic_variable, scales = "free_y") +
  xlab("Day of the Week") + ylab("Hourly Average") +
  theme_tufte()

# we print the graph
graph_road_traffic_wd

# save the graph
ggsave(
  graph_road_traffic_wd,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "2.road_traffic",
    "graph_road_traffic_wd.pdf"
  ),
  width = 30,
  height = 15,
  units = "cm",
  device = cairo_pdf
)
```

We then plot the empirical cumulative distribution of road traffic by weekdays and weekend:

```{r, fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
graph_ecdf_road_traffic_we <- data %>%
  select(weekend, road_traffic_flow_all, road_occupancy_rate) %>%
  pivot_longer(cols = -c(weekend), names_to = "traffic_variable", values_to = "value") %>%
  mutate(traffic_variable = ifelse(traffic_variable == "road_traffic_flow_all", "Flow of Vehicles (N)", "Occupancy Rate (%)")) %>%
  mutate(weekend = ifelse(weekend == 1, "Weekend", "Weekday")) %>%
  ggplot(., aes(x = value, colour = weekend)) +
  stat_ecdf(size = 1.1) +
  scale_color_manual(values = c(my_blue, my_orange)) +
  facet_wrap(~ traffic_variable, scales = "free_x") +
  ylab("Cumulative Probability") + xlab("Road Traffic Value") +
  labs(colour = "Status:") +
  theme_tufte()

# we print the graph
graph_ecdf_road_traffic_we

# save the graph
ggsave(
  graph_ecdf_road_traffic_we,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "2.road_traffic",
    "graph_ecdf_road_traffic_we.pdf"
  ),
  width = 30,
  height = 18,
  units = "cm",
  device = cairo_pdf
)
```

### NO2 Variation by Day of the Week

We plot the average NO2 concentration by day of the week:

```{r, fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
# make the graph
graph_no2_wd <- data %>%
  select(weekday, mean_no2_l, mean_no2_sl) %>%
  rename("Longchamp Station" = mean_no2_l,
         "Saint-Louis Station" = mean_no2_sl) %>%
  pivot_longer(cols = -c(weekday),
               names_to = "station",
               values_to = "concentration") %>%
  group_by(weekday, station) %>%
  summarise(mean_no2 = mean(concentration, na.rm = TRUE)) %>%
  ggplot(., aes(x = weekday, y = mean_no2)) +
  geom_segment(aes(
    x = weekday,
    xend = weekday,
    y = 0,
    yend = mean_no2
  )) +
  geom_point(
    shape = 21,
    size = 4,
    colour = "black",
    fill = my_blue
  ) +
  facet_wrap( ~ station, ncol = 1) +
  xlab("Day of the Week") + ylab("Average NO2 Concentrations (µg/m³)") +
  theme_tufte()

# we print the graph
graph_no2_wd

# save the graph
ggsave(
  graph_no2_wd,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "2.road_traffic",
    "graph_no2_wd.pdf"
  ),
  width = 30,
  height = 25,
  units = "cm",
  device = cairo_pdf
)
```


We also plot the empirical cumulative distribution of NO$_{2}$ concentration by weekdays and weekend:

```{r, fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
# make the graph
graph_ecdf_no2_we <- data %>%
  select(weekend, mean_no2_l, mean_no2_sl) %>%
  mutate(weekend = ifelse(weekend == 1, "Weekend", "Weekday")) %>%
  rename("Longchamp Station" = mean_no2_l,
         "Saint-Louis Station" = mean_no2_sl) %>%
  pivot_longer(cols = -c(weekend),
               names_to = "station",
               values_to = "concentration") %>%
  ggplot(., aes(x = concentration, colour = weekend)) +
  stat_ecdf(size = 1.1) +
  scale_color_manual(values = c(my_blue, my_orange)) +
  facet_wrap( ~ station) +
  ylab("Cumulative Probability") + xlab("NO2 Concentrations (µg/m³)") +
  labs(colour = "Status:") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# we print the graph
graph_ecdf_no2_we

# save the graph
ggsave(
  graph_ecdf_no2_we,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "2.road_traffic",
    "graph_ecdf_no2_we.pdf"
  ),
  width = 30,
  height = 15,
  units = "cm",
  device = cairo_pdf
)
```

### Average Total Gross Tonnage of Cruise Vessels by Day of the Week

We plot the average total gross tonnage of vessels by day of the week:

```{r, fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
graph_gross_tonnage_cruise_wd <- data %>%
  select(weekday, total_gross_tonnage_cruise) %>%
  group_by(weekday) %>%
  summarise(total_gross_tonnage_cruise = mean(total_gross_tonnage_cruise, na.rm = TRUE)) %>%
  ggplot(., aes(x = weekday, y = total_gross_tonnage_cruise)) +
  geom_segment(aes(
    x = weekday,
    xend = weekday,
    y = 0,
    yend = total_gross_tonnage_cruise
  )) +
  geom_point(
    shape = 21,
    size = 4,
    colour = "black",
    fill = my_blue
  ) +
  scale_y_continuous(
    breaks = scales::pretty_breaks(n = 5),
    labels = function(x)
      format(x, big.mark = " ", scientific = FALSE)
  ) +
  xlab("Day of the Week") + ylab("Average Total Gross Tonnage") +
  theme_tufte()

# we print the graph
graph_gross_tonnage_cruise_wd

# save the graph
ggsave(
  graph_gross_tonnage_cruise_wd,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "2.road_traffic",
    "graph_gross_tonnage_cruise_wd.pdf"
  ),
  width = 30,
  height = 15,
  units = "cm",
  device = cairo_pdf
)
```

# Weather Covariates Balance

We can check that the weekends and weekdays are similar for weather covariates. For continuous weather covariates, we draw boxplots for treated and control groups:

```{r, fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
# we select control variables and store them in a long dataframe
data_weather_continuous_variables <- data %>%
  mutate(weekend = ifelse(weekend == 1, "Weekend", "Weekday")) %>%
  select(temperature_average,
         humidity_average,
         wind_speed,
         weekend) %>%
  pivot_longer(cols = -c(weekend),
               names_to = "variable",
               values_to = "values") %>%
  mutate(
    variable = case_when(
      variable == "temperature_average" ~ "Average Temperature (°C)",
      variable == "humidity_average" ~ "Humidity Average (%)",
      variable == "wind_speed" ~ "Wind Speed (m/s)"
    )
  )


graph_boxplot_continuous_weather <-
  ggplot(data_weather_continuous_variables,
         aes(x = weekend, y = values, colour = weekend)) +
  geom_violin(size = 1.2) +
  geom_boxplot(width = 0.1, outlier.shape = NA) +
  scale_color_manual(values = c(my_blue, my_orange)) +
  ylab("Covariate Value") +
  xlab("") +
  labs(colour = "Units Status:") +
  facet_wrap( ~ variable, scale = "free", ncol = 3) +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# we print the graph
graph_boxplot_continuous_weather

# save the graph
ggsave(
  graph_boxplot_continuous_weather,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "2.road_traffic",
    "graph_boxplot_continuous_weather.pdf"
  ),
  width = 40,
  height = 15,
  units = "cm",
  device = cairo_pdf
)
```

For the rainfall dummy and the wind direction categories, we plot the proportions:

```{r, fig.width=10, fig.height=5, layout="l-page", code_folding="Please show me the code!"}
# we select the rainfall variables
data_weather_categorical <- data %>%
  mutate(weekend = ifelse(weekend == 1, "Weekend", "Weekday")) %>%
  select(rainfall_height_dummy,
         wind_direction_categories,
         weekend) %>%
  drop_na() %>%
  mutate_at(vars(rainfall_height_dummy), ~ ifelse(. == 1, "True", "False")) %>%
  mutate_all(~ as.character(.)) %>%
  pivot_longer(cols = -c(weekend),
               names_to = "variable",
               values_to = "values") %>%
  # group by weekend, variable and values
  group_by(weekend, variable, values) %>%
  # compute the number of observations
  summarise(n = n()) %>%
  # compute the proportion
  mutate(freq = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(
    variable = case_when(
      variable == "wind_direction_categories" ~ "Wind Direction",
      variable == "rainfall_height_dummy" ~ "Rainfall Dummy"
    )
  )

# build the graph
graph_categorical_weather <-
  ggplot(data_weather_categorical, aes(x = freq, y = values, fill = weekend)) +
  geom_point(shape = 21,
             size = 6,
             alpha = 0.8) +
  scale_fill_manual(values = c(my_blue, my_orange)) +
  facet_wrap(~ variable, scales = "free") +
  xlab("Proportion (%)") +
  ylab("") +
  labs(fill = "Units Status:") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# we print the graph
graph_categorical_weather

# save the graph
ggsave(
  graph_categorical_weather,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "2.road_traffic",
    "graph_categorical_weather.pdf"
  ),
  width = 40,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

# Weekend Effect on Road Traffic and NO$_{2}$ Concentrations

We calculate the average differences in road traffic and NO$_{2}$ concentrations between weekends and weekdays.

### Weekend Effect on Road Traffic

We compute the average difference in the flow of vehicles and road occupancy rate between weekdays and weekends:

```{r}
# compute difference in flow
diff_flow <- data %>%
  filter(!is.na(road_traffic_flow_all)) %>%
  summarise(average_difference = mean(road_traffic_flow_all[weekend == 1]) - mean(road_traffic_flow_all[weekend == 0]))

# compute difference in occupancy rate
diff_occupancy <- data %>%
  filter(!is.na(road_occupancy_rate)) %>%
  summarise(average_difference = mean(road_occupancy_rate[weekend == 1]) - mean(road_occupancy_rate[weekend == 0]))
```

On average, the hourly road traffic decreases by `r round(abs(diff_flow),0)` vehicles on weekends. There is a drop of `r round(diff_occupancy, 1)` poins in the occupancy rate.

### Weekend Effect on NO2 concentrations

We compute the average effect of weekends on NO2 concentrations.

```{r}
# compute mean difference in no2 for the two stations
diff_no2 <- data %>%
  select(weekend, mean_no2_l, mean_no2_sl) %>%
  pivot_longer(cols = -c(weekend),
               names_to = "station",
               values_to = "concentration") %>%
  group_by(station) %>%
  summarise(average_difference = round(mean(concentration[weekend == 1], na.rm = TRUE) - mean(concentration[weekend == 0], na.rm = TRUE), 1))
```

On average, NO$_{2}$ is lower between `r abs(diff_no2[1,2])` and `r abs(diff_no2[2,2])` $\mu g/m^{3}$ on weekends compared to weekdays, depending on the monitoring station.







