---
title: "Checking Balance Improvement"
description: |
  Comparing days with cruise traffic to days without. Adjusting for calendar and weather indicators.
author:
  - name: Marie-Abèle Bind 
    url: https://scholar.harvard.edu/marie-abele
    affiliation: Biostatistics Center, Massachusetts General Hospital
    affiliation_url: https://biostatistics.massgeneral.org/faculty/marie-abele-bind-phd/
  - name: Marion Leroutier 
    url: https://www.parisschoolofeconomics.eu/en/leroutier-marion/work-in-progress/
    affiliation: Misum, Stockholm School of Economics
    affiliation_url: https://www.hhs.se/en/persons/l/leroutier-marion/
  - name: Léo Zabrocki 
    url: https://lzabrocki.github.io/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/fr/zabrocki-leo/
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
      toc: true
      toc_depth: 3
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
# code chunk option
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  dev = "CairoPNG",
  dpi = 600
)
```

In this document, we take great care providing all steps and R codes required to check whether our matching procedure allowed to improve covariates balance. We compare days where:

* treated units are days with positive cruise traffic in *t*.
* control units are days without cruise traffic in *t*.

We adjust for calendar calendar indicator and weather confounding factors.

**Should you have any questions, need help to reproduce the analysis or find coding errors, please do not hesitate to contact us at leo.zabrocki@psemail.eu and marion.leroutier@hhs.se.**

# Required Packages

To reproduce exactly the `2_7_daily_checking_balance_improvement.html` document, we first need to have installed:

* the [R](https://www.r-project.org/) programming language 
* [RStudio](https://rstudio.com/), an integrated development environment for R, which will allow you to knit the `2_7_daily_checking_balance_improvement.Rmd` file and interact with the R code chunks
* the [R Markdown](https://rmarkdown.rstudio.com/) package
* and the [Distill](https://rstudio.github.io/distill/) package which provides the template for this document. 

Once everything is set up, we have to load the following packages:

```{r}
# load required packages
library(knitr) # for creating the R Markdown document
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualization
library(dtplyr) # to speed up to dplyr
library(randChecks) # for randomization check
library(ggridges) # for ridge density plots
library(Cairo) # for printing custom police of graphs
library(patchwork) # combining plots
```

We finally load our custom `ggplot2` theme for graphs:

```{r}
# load ggplot custom theme
source(here::here("inputs",
                  "2.functions",
                  "script_theme_tufte.R"))
# define nice colors
my_blue <- "#0081a7"
my_orange <- "#fb8500"
```


# Preparing the Data

We load the initial and matched data and bind them together:

```{r}
# load matching data
data_matching <-
  readRDS(
    here::here(
      "inputs",
      "1.data",
      "2.daily_data",
      "2.data_for_analysis",
      "1.cruise_experiment",
      "matching_data.rds"
    )
  ) %>%
  mutate(dataset = "Initial Data")

# load matched data
data_matched <-
  readRDS(
    here::here(
      "inputs",
      "1.data",
      "2.daily_data",
      "2.data_for_analysis",
      "1.cruise_experiment",
      "matched_data.rds"
    )
  ) %>%
  mutate(dataset = "Matched Data")

# bind the two datasets
data <- bind_rows(data_matching, data_matched)
```

We change labels of the `is_treated` variable :

```{r}
data <- data %>%
  mutate(is_treated = ifelse(is_treated == "TRUE", "True", "False"))
```

# Love Plots

### Continuous Weather Covariates

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# compute figures for the love plot
data_weather_continuous <- data %>%
  dplyr::select(
    dataset,
    is_treated,
    contains("temperature"),
    contains("humidity"),
    contains("wind_speed")
  ) %>%
  pivot_longer(
    cols = -c(is_treated, dataset),
    names_to = "variable",
    values_to = "values"
  ) %>%
  mutate(
    weather_variable = NA %>%
      ifelse(
        str_detect(variable, "temperature_average"),
        "Average Temperature",
        .
      ) %>%
      ifelse(
        str_detect(variable, "humidity_average"),
        "Humidity Average",
        .
      ) %>%
      ifelse(str_detect(variable, "wind_speed"), "Wind Speed", .)
  ) %>%
  mutate(time = "0" %>%
           ifelse(str_detect(variable, "lag_1"), "-1", .) %>%
           ifelse(str_detect(variable, "lead_1"), "+1", .)) %>%
  mutate(time = fct_relevel(time, "-1", "0", "+1")) %>%
  dplyr::select(dataset, is_treated, weather_variable, time, values)

data_abs_difference_continuous_weather <-
  data_weather_continuous %>%
  group_by(dataset, weather_variable, time, is_treated) %>%
  summarise(mean_values = mean(values, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_values[2] - mean_values[1]))

data_sd_weather_continuous <-  data_weather_continuous %>%
  filter(dataset == "Initial Data" & is_treated == "True") %>%
  group_by(weather_variable, time, is_treated) %>%
  summarise(sd_treatment = sd(values, na.rm = TRUE)) %>%
  ungroup() %>%
  select(weather_variable, time, sd_treatment)

data_love_continuous_weather <-
  left_join(
    data_abs_difference_continuous_weather,
    data_sd_weather_continuous,
    by = c("weather_variable", "time")
  ) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment))

# make the graph
graph_love_plot_continuous_weather <-
  ggplot(
    data_love_continuous_weather,
    aes(
      y = fct_rev(time),
      x = standardized_difference,
      colour = fct_rev(dataset),
      shape = fct_rev(dataset)
    )
  ) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  facet_wrap( ~ weather_variable, scales = "free_y") +
  xlab("Standardized Mean Differences") +
  ylab("Day") +
  theme_tufte()

# plot the graph
graph_love_plot_continuous_weather

# save the graph
ggsave(
  graph_love_plot_continuous_weather,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_continuous_weather.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

### Categorical Weather Covariates

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# compute figures for the love plot
data_weather_categorical <- data %>%
  select(
    dataset,
    is_treated,
    contains("rainfall_height_dummy"),
    contains("wind_direction_east_west")
  ) %>%
  mutate_at(vars(contains("rainfall")), ~ ifelse(. == 1, "True", "False")) %>%
  mutate_all( ~ as.character(.)) %>%
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "variable",
    values_to = "values"
  ) %>%
  # group by is_treated, variable and values
  group_by(dataset, is_treated, variable, values) %>%
  # compute the number of observations
  summarise(n = n()) %>%
  # compute the proportion
  mutate(freq = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(
    weather_variable = NA %>%
      ifelse(str_detect(variable, "wind"), "Wind Direction", .) %>%
      ifelse(str_detect(variable, "rainfall"), "Rainfall Dummy", .)
  ) %>%
  mutate(time = "t" %>%
           ifelse(str_detect(variable, "lag_1"), "t-1", .) %>%
           ifelse(str_detect(variable, "lead_1"), "t+1", .)) %>%
  filter(!is.na(time)) %>%
  mutate(variable = paste(weather_variable, time, sep = " ")) %>%
  select(dataset, is_treated, weather_variable, variable, values, freq) %>%
  pivot_wider(names_from = is_treated, values_from = freq) %>%
  mutate(abs_difference = abs(`True` - `False`)) %>%
  filter(values != "False")

# create the figure for wind direction
graph_love_plot_wind_direction <- data_weather_categorical %>%
  filter(weather_variable == "Wind Direction") %>%
  mutate(variable = fct_relevel(
    variable,
    "Wind Direction t-1",
    "Wind Direction t",
    "Wind Direction t+1"
  )) %>%
  ggplot(.,
         aes(
           y = fct_rev(values),
           x = abs_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  facet_wrap( ~ variable, scales = "free_y", ncol = 3) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte()

# print the figure for wind direction
graph_love_plot_wind_direction

# save the figure for wind direction
ggsave(
  graph_love_plot_wind_direction,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_wind_direction.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)

# create the figure for rainfall dummy
graph_love_plot_rainfall <- data_weather_categorical %>%
  filter(weather_variable == "Rainfall Dummy") %>%
  mutate(variable = fct_relevel(
    variable,
    "Rainfall Dummy t-1",
    "Rainfall Dummy t",
    "Rainfall Dummy t+1"
  )) %>%
  ggplot(.,
         aes(
           y = fct_rev(variable),
           x = abs_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte()

# print the figure for rainfall dummy
graph_love_plot_rainfall

# save the figure for rainfall dummy
ggsave(
  graph_love_plot_rainfall,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_rainfall.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

### Pollutants

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# compute figures for the love plot
data_pollutants <- data %>%
  select(
    dataset,
    is_treated,
    contains("no2"),
    contains("o3"),
    contains("pm10"),
    contains("pm25"),
    contains("so2")
  ) %>%
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "variable",
    values_to = "values"
  ) %>%
  mutate(
    pollutant = NA %>%
      ifelse(str_detect(variable, "no2_l"), "NO2 Longchamp", .) %>%
      ifelse(str_detect(variable, "no2_sl"), "NO2 Saint-Louis", .) %>%
      ifelse(str_detect(variable, "o3"), "O3 Lonchamp", .) %>%
      ifelse(str_detect(variable, "pm10_l"), "PM10 Longchamp", .) %>%
      ifelse(str_detect(variable, "pm10_sl"), "PM10 Saint-Louis", .) %>%
      ifelse(str_detect(variable, "pm25"), "PM2.5 Longchamp", .) %>%
      ifelse(str_detect(variable, "so2"), "SO2 Longchamp", .)
  ) %>%
  mutate(time = NA %>%
           ifelse(str_detect(variable, "lag_1"), "-1", .)) %>%
  filter(!is.na(time)) %>%
  select(dataset, is_treated, pollutant, time, values)

data_abs_difference_pollutants <- data_pollutants %>%
  group_by(dataset, pollutant, time, is_treated) %>%
  summarise(mean_values = mean(values, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_values[2] - mean_values[1]))

data_sd_pollutants <-  data_pollutants %>%
  filter(dataset == "Initial Data" & is_treated == "True") %>%
  group_by(pollutant, time, is_treated) %>%
  summarise(sd_treatment = sd(values, na.rm = TRUE)) %>%
  ungroup() %>%
  select(pollutant, time, sd_treatment)

data_love_pollutants <-
  left_join(
    data_abs_difference_pollutants,
    data_sd_pollutants,
    by = c("pollutant", "time")
  ) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment))

# create the graph
graph_love_plot_pollutants <-
  ggplot(
    data_love_pollutants,
    aes(
      y = fct_rev(time),
      x = standardized_difference,
      colour = fct_rev(dataset),
      shape = fct_rev(dataset)
    )
  ) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  facet_wrap( ~ pollutant, ncol = 4) +
  xlab("Standardized Mean Differences") +
  ylab("Day") +
  theme_tufte()
# print the graph
graph_love_plot_pollutants

# save the graph
ggsave(
  graph_love_plot_pollutants,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_pollutants.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

### Vessels Traffic

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# compute figures for the love plot
data_tonnage <-  data %>%
  # select relevant variables
  select(
    dataset,
    is_treated,
    contains("total_gross_tonnage_cruise"),
    contains("total_gross_tonnage_ferry"),
    contains("total_gross_tonnage_other_boat")
  ) %>%
  # transform data in long format
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "variable",
    values_to = "tonnage"
  ) %>%
  # create vessel type variable
  mutate(
    vessel_type = NA %>%
      ifelse(str_detect(variable, "cruise"), "Cruise", .) %>%
      ifelse(str_detect(variable, "ferry"), "Ferry", .) %>%
      ifelse(str_detect(variable, "other_boat"), "Other Types of Vessels", .)
  ) %>%
  mutate(time = NA %>%
           ifelse(str_detect(variable, "lag_1"), "-1", .) %>%
           ifelse(str_detect(variable, "lead_1"), "1", .)) %>%
  filter(!is.na(time)) %>%
  select(dataset, vessel_type, is_treated, time, tonnage)

data_abs_difference_tonnage <- data_tonnage %>%
  group_by(dataset, vessel_type, time, is_treated) %>%
  summarise(mean_tonnage = mean(tonnage, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_tonnage[2] - mean_tonnage[1]))

data_sd_tonnage <-  data_tonnage %>%
  filter(dataset == "Initial Data" & is_treated == "True") %>%
  group_by(vessel_type, time, is_treated) %>%
  summarise(sd_treatment = sd(tonnage, na.rm = TRUE)) %>%
  ungroup() %>%
  select(vessel_type, time, sd_treatment)

data_love_tonnage <-
  left_join(data_abs_difference_tonnage,
            data_sd_tonnage,
            by = c("vessel_type", "time")) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment)) %>%
  filter(!(vessel_type == "Cruise" & time == 0))

# create the graph
graph_love_plot_tonnage <-
  ggplot(
    data_love_tonnage,
    aes(
      x = standardized_difference,
      y = as.factor(time),
      colour = fct_rev(dataset),
      shape = fct_rev(dataset)
    )
  ) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +  facet_wrap( ~ vessel_type) +
  xlab("Standardized Mean Differences") +
  ylab("Day") +
  theme_tufte()
# print the graph
graph_love_plot_tonnage

# save the graph
ggsave(
  graph_love_plot_tonnage,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_tonnage.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```


### Road Traffic

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# compute figures for the love plot
data_road <-  data %>%
  # select relevant variables
  select(dataset, is_treated, contains("road_traffic_flow_all"), contains("road_occupancy_rate")) %>%
  # transform data in long format
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "road_traffic",
    values_to = "value"
  ) %>%
  mutate(time = 0 %>%
           ifelse(str_detect(road_traffic, "lag_1"),-1, .) %>%
           ifelse(str_detect(road_traffic, "lead_1"), 1, .)) %>%
  mutate(road_traffic = ifelse(str_detect(road_traffic, "road_traffic_flow_all"), "Flow of Vehicles", "Occupany Rate")) %>%
  select(dataset, road_traffic, is_treated, time, value)

data_abs_difference_road <- data_road %>%
  group_by(dataset, road_traffic, time, is_treated) %>%
  summarise(mean_value = mean(value, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_value[2] - mean_value[1]))

data_sd_tonnage <-  data_road %>%
  filter(dataset == "Initial Data" & is_treated == "True") %>%
  group_by(road_traffic, time, is_treated) %>%
  summarise(sd_treatment = sd(value, na.rm = TRUE)) %>%
  ungroup() %>%
  select(road_traffic, time, sd_treatment)

data_love_road <-
  left_join(data_abs_difference_road,
            data_sd_tonnage,
            by = c("road_traffic", "time")) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment))


# create the graph
graph_love_plot_road <-
  ggplot(
    data_love_road,
    aes(
      x = standardized_difference,
      y = as.factor(time),
      colour = fct_rev(dataset),
      shape = fct_rev(dataset)
    )
  ) +
  geom_vline(xintercept = 0) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  facet_wrap(~ road_traffic) + 
  xlab("Standardized Mean Differences") +
  ylab("Day") +
  theme_tufte()

# print the graph
graph_love_plot_road

# save the graph
ggsave(
  graph_love_plot_road,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_road.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

### Calendar Indicators

Create the relevant data:

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# compute figures for the love plot
data_calendar <- data %>%
  mutate(weekday = lubridate::wday(date, abbr = FALSE, label = TRUE)) %>%
  select(dataset,
         is_treated,
         weekday,
         holidays_dummy,
         bank_day_dummy,
         month,
         year) %>%
  mutate_at(vars(holidays_dummy, bank_day_dummy),
            ~ ifelse(. == 1, "True", "False")) %>%
  mutate_all( ~ as.character(.)) %>%
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "variable",
    values_to = "values"
  ) %>%
  # group by is_treated, variable and values
  group_by(dataset, is_treated, variable, values) %>%
  # compute the number of observations
  summarise(n = n()) %>%
  # compute the proportion
  mutate(freq = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(
    calendar_variable = NA %>%
      ifelse(str_detect(variable, "weekday"), "Day of the Week", .) %>%
      ifelse(str_detect(variable, "holidays_dummy"), "Holidays", .) %>%
      ifelse(str_detect(variable, "bank_day_dummy"), "Bank Day", .) %>%
      ifelse(str_detect(variable, "month"), "Month", .) %>%
      ifelse(str_detect(variable, "year"), "Year", .)
  ) %>%
  select(dataset, is_treated, calendar_variable, values, freq) %>%
  pivot_wider(names_from = is_treated, values_from = freq) %>%
  mutate(abs_difference = abs(`True` - `False`)) %>%
  filter(values != "False")
```

Plot for bank days and holidays:

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# graph for bank days and holidays
graph_love_plot_bank_holidays <- data_calendar %>%
  filter(calendar_variable %in% c("Bank Day", "Holidays")) %>%
  ggplot(.,
         aes(
           y = values,
           x = abs_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  facet_wrap(~ calendar_variable) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte()
# print the plot
graph_love_plot_bank_holidays

# save the plot
ggsave(
  graph_love_plot_bank_holidays,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_bank_holidays.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

Plot for days of the week:

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# graph for weekdays
graph_love_plot_weekday <- data_calendar %>%
  filter(calendar_variable == "Day of the Week") %>%
  mutate(
    values = fct_relevel(
      values,
      "Monday",
      "Tuesday",
      "Wednesday",
      "Thursday",
      "Friday",
      "Saturday",
      "Sunday"
    )
  ) %>%
  ggplot(., aes(
    y = fct_rev(values),
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# print the plot
graph_love_plot_weekday

# save the plot
ggsave(
  graph_love_plot_weekday,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_weekday.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

Plot for months:

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# graph for month
graph_love_plot_month <- data_calendar %>%
  filter(calendar_variable == "Month") %>%
  mutate(
    values = fct_relevel(
      values,
      "January",
      "February",
      "March",
      "April",
      "May",
      "June",
      "July",
      "August",
      "September",
      "October",
      "November",
      "December"
    )
  ) %>%
  ggplot(., aes(
    y = fct_rev(values),
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  ggtitle("Month") +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# print the plot
graph_love_plot_month

# save the plot
ggsave(
  graph_love_plot_month,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_month.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

Plot for years:

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# graph for year
graph_love_plot_year <- data_calendar %>%
  filter(calendar_variable == "Year") %>%
  ggplot(., aes(
    y = as.factor(as.numeric(values)),
    x = abs_difference,
    colour = fct_rev(dataset),
    shape = fct_rev(dataset)
  )) +
  geom_vline(xintercept = 0) +
  geom_point(size = 2, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_orange, my_blue)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  ggtitle("Year") +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# print the graph
graph_love_plot_year

# save the plot
ggsave(
  graph_love_plot_year,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_love_plot_year.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

# Overall Balance Improvement

We plot the distribution of standardized mean differences for continuous covariates or the absolute percentage points differences for categorical covariates between treated and control units before and after matching.

### Continuous Covariates

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# we select the dataset indicator and the standardized difference
data_love_tonnage <- data_love_tonnage %>%
  ungroup() %>%
  filter(time < 0) %>%
  select(dataset, standardized_difference)

data_love_pollutants <- data_love_pollutants %>%
  ungroup() %>%
  filter(time < 0) %>%
  select(dataset, standardized_difference)

data_love_continuous_weather <- data_love_continuous_weather %>%
  ungroup() %>%
  filter(time <= 0) %>%
  select(dataset, standardized_difference)

data_love_road <- data_love_road %>%
  ungroup() %>%
  filter(time <= 0) %>%
  select(dataset, standardized_difference)

data_continuous_love <-
  bind_rows(data_love_tonnage, data_love_pollutants) %>%
  bind_rows(., data_love_continuous_weather) %>%
  bind_rows(., data_love_road)

# create the graph
graph_boxplot_continuous_balance_improvement <-
  ggplot(data_continuous_love,
         aes(x = dataset, y = standardized_difference)) +
  geom_boxplot(colour = my_blue) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  xlab("Dataset") +
  ylab("Standardized Mean Differences") +
  theme_tufte()
# print the graph
graph_boxplot_continuous_balance_improvement

# save the graph
ggsave(
  graph_boxplot_continuous_balance_improvement,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_boxplot_continuous_balance_improvement.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

### Categorical Covariates

```{r fig.width=10, fig.height=5, code_folding="Please show me the code!"}
# we select the dataset indicator and the standardized difference
data_calendar <- data_calendar %>%
  ungroup() %>%
  select(dataset, abs_difference)

data_weather_categorical <- data_weather_categorical %>%
  ungroup() %>%
  filter(
    variable %in% c(
      "Rainfall Dummy t",
      "Rainfall Dummy t-1",
      "Wind Direction t",
      "Wind Direction t-1"
    )
  ) %>%
  select(dataset, abs_difference)

data_categorical_love <-
  bind_rows(data_calendar, data_weather_categorical)

# create the graph
graph_boxplot_categorical_balance_improvement <-
  ggplot(data_categorical_love, aes(x = dataset, y = abs_difference)) +
  geom_boxplot(colour = my_blue) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  xlab("Dataset") +
  ylab("Absolute Difference \nin Percentage Points") +
  theme_tufte()

# print the graph
graph_boxplot_categorical_balance_improvement

# save the graph
ggsave(
  graph_boxplot_categorical_balance_improvement,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_boxplot_categorical_balance_improvement.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```


We combine the two previous plots in a single figure:

```{r, code_folding="Please show me the code!"}
# combine plots
graph_overall_balance <-
  graph_boxplot_continuous_balance_improvement /graph_boxplot_categorical_balance_improvement +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 30, face = "bold"))

# save the plot
ggsave(
  graph_overall_balance,
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "graph_overall_balance.pdf"
  ),
  width = 20,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
```

# Randomization Check for Covariate Balance

Finally, we implement the procedure of Zach Branson (2021) to carry out a randomization check to test whether daily observations are as-if randomized according to observed covariates.

```{r, eval = FALSE}
# we select covariates
data_covs <- data_matched %>%
  select(
    pair_number,
    is_treated,
    year,
    month,
    holidays_dummy,
    bank_day_dummy,
    total_gross_tonnage_ferry:total_gross_tonnage_other_boat,
    temperature_average:wind_speed,
    wind_direction_east_west,
    holidays_dummy_lag_1,
    bank_day_dummy_lag_1,
    total_gross_tonnage_cruise_lag_1:total_gross_tonnage_other_boat_lag_1,
    temperature_average_lag_1:wind_speed_lag_1,
    wind_direction_east_west_lag_1,
    mean_no2_l_lead_1:mean_o3_l_lead_1
  )

# recode some variables
data_covs <- data_covs %>%
  mutate(is_treated = ifelse(is_treated == TRUE, 1, 0)) %>%
  mutate_at(
    vars(wind_direction_east_west , wind_direction_east_west_lag_1),
    ~ ifelse(. == "West", 1, 0)
  ) %>%
  fastDummies::dummy_cols(., select_columns = c("year", "month")) %>%
  select(-c("year", "month"))

# format data for asIfRandPlot() function
pair_number <- as.numeric(data_covs$pair_number)
is_treated <- data_covs$is_treated

data_covs <- data_covs %>%
  select(-pair_number,-is_treated) %>%
  select(-year_2018, - month_December, -bank_day_dummy, - bank_day_dummy_lag_1) 

# run balance check
asIfRandPlot(
  X.matched = data_covs,
  indicator.matched = is_treated,
  assignment = c("complete", "blocked"),
  subclass = pair_number,
  perms = 100
)
```

```{r, echo = FALSE, out.width = "500%", fig.align = "center"}
knitr::include_graphics(
    here::here("inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "1.checking_matching_procedure",
    "ri_balance_check.png"))
```
















