---
title: "Exploraty Data Analysis of the Effect of Road Traffic on NO2 Concentration in Marseille"
description: |
  Comparing weekdays to weekend.
author:
  - name: Marie-Abèle Bind 
    url: https://scholar.harvard.edu/marie-abele
    affiliation: Biostatistics Center, Massachusetts General Hospital
    affiliation_url: https://biostatistics.massgeneral.org/faculty/marie-abele-bind-phd/
  - name: Marion Leroutier 
    url: https://www.parisschoolofeconomics.eu/en/leroutier-marion/work-in-progress/
    affiliation: Misum, Stockholm School of Economics
    affiliation_url: https://www.hhs.se/en/persons/l/leroutier-marion/
  - name: Léo Zabrocki 
    url: https://lzabrocki.github.io/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/fr/zabrocki-leo/
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
      toc: true
      toc_depth: 3
editor_options: 
  chunk_output_type: console
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
# code chunk option
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  dev = "CairoPNG",
  dpi = 300
)
```

In this document

**Should you have any questions, need help to reproduce the analysis or find coding errors, please do not hesitate to contact us at leo.zabrocki@psemail.eu and marion.leroutier@hhs.se.**

# Required Packages

To reproduce exactly the `2_9_daily_regression_analysis.html` document, we first need to have installed:

* the [R](https://www.r-project.org/) programming language on your computer 
* [RStudio](https://rstudio.com/), an integrated development environment for R, which will allow you to knit the `2_9_daily_regression_analysis.Rmd` file and interact with the R code chunks
* the [R Markdown](https://rmarkdown.rstudio.com/) package
* and the [Distill](https://rstudio.github.io/distill/) package which provides the template of this document. 

Once everything is set up, we have to load the following packages:

```{r}
# load required packages
library(knitr) # for creating the R Markdown document
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualization
library(broom) # for cleaning regression outputs
library(randChecks) # for randomization check
library(lmtest) # for modifying regression standard errors
library(sandwich) # for robust and cluster robust standard errors
library(kableExtra) # for table formatting
library(Cairo) # for printing custom police of graphs
```

We load our custom `ggplot2` theme for graphs:

```{r}
# load ggplot custom theme
source(here::here("inputs",
                  "2.functions",
                  "script_theme_tufte.R"))
# define nice colors
my_blue <- "#0081a7"
my_orange <- "#fb8500"
```

```{r}
# load matching data
data_matching <-
  readRDS(
    here::here(
      "inputs",
      "1.data",
      "2.daily_data",
      "2.data_for_analysis",
      "1.cruise_experiment",
      "matching_data.rds"
    )
  ) %>%
  mutate(year = as.factor(year)) %>%
  select(
    is_treated,
    year,
    month,
    weekday,
    holidays_dummy,
    bank_day_dummy,
    total_gross_tonnage_ferry:total_gross_tonnage_other_boat,
    temperature_average:wind_speed,
    wind_direction_east_west,
    holidays_dummy_lag_1,
    bank_day_dummy_lag_1,
    total_gross_tonnage_cruise_lag_1:total_gross_tonnage_other_boat_lag_1,
    temperature_average_lag_1:wind_speed_lag_1,
    wind_direction_east_west_lag_1,
    contains("no2"),
    contains("o3"),
    contains("pm10"),
    contains("pm25"),
    contains("so2")
  )
```



# Balance check

```{r, eval = FALSE}
# we select covariates
data_covs <- data_matching %>%
  select(
    is_treated,
    year,
    month,
    weekday,
    holidays_dummy,
    bank_day_dummy,
    total_gross_tonnage_ferry:total_gross_tonnage_other_boat,
    temperature_average:wind_speed,
    wind_direction_east_west,
    holidays_dummy_lag_1,
    bank_day_dummy_lag_1,
    total_gross_tonnage_cruise_lag_1:total_gross_tonnage_other_boat_lag_1,
    temperature_average_lag_1:wind_speed_lag_1,
    wind_direction_east_west_lag_1,
    mean_no2_l_lag_1:mean_o3_l_lag_1
  ) %>%
  drop_na()

data_covs <- data_covs %>%
  group_by(year, month) %>%
  mutate(year_month_block = cur_group_id())

# recode some variables
data_covs <- data_covs %>%
  mutate(is_treated = ifelse(is_treated == TRUE, 1, 0)) %>%
  mutate_at(
    vars(wind_direction_east_west , wind_direction_east_west_lag_1),
    ~ ifelse(. == "West", 1, 0)
  ) %>%
  fastDummies::dummy_cols(., select_columns = c("year", "month", "weekday")) %>%
  select(-c("year", "month", "weekday"))

# format data for asIfRandPlot() function
is_treated <- data_covs$is_treated
year_month_block <- data_covs$year_month_block

data_covs <- data_covs %>%
  select(-is_treated, -year_month_block) %>%
  select(-year_2018,-month_December, -weekday_Sunday)

# run balance check
asIfRandPlot(
  X.matched = data_covs,
  indicator.matched = is_treated,
  assignment = c("complete", "blocked"),
  subclass = year_month_block,
  perms = 1000
)
```


```{r, echo = FALSE, out.width = "800%", fig.align = "center"}
knitr::include_graphics(
  here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "3.outcome_regression",
    "ri_balance_check.png"
  )
)
```


# Regression Analysis



```{r}
function_reg_analysis <- function(data) {
  # we fit the regression model
  model <-
    lm(
      concentration ~ is_treated + total_gross_tonnage_ferry + total_gross_tonnage_other_boat +
        total_gross_tonnage_ferry_lag_1 + total_gross_tonnage_other_boat_lag_1 +
        temperature_average + temperature_average_lag_1 +
        humidity_average + humidity_average_lag_1 +
        rainfall_height_dummy + rainfall_height_dummy_lag_1 +
        wind_direction_east_west + wind_direction_east_west_lag_1 +
        wind_speed + wind_speed_lag_1 +
        weekday + holidays_dummy + bank_day_dummy + month * year,
      data = data
    )
  
  # retrieve the estimate and 95% ci
  results_reg <- tidy(coeftest(model, vcov = vcovHC),
                      conf.int = TRUE) %>%
    filter(term == "is_treatedTRUE") %>%
    select(term, estimate, conf.low, conf.high)
  
  # return output
  return(results_reg)
}


# reshape in long according to pollutants
data_reg_analysis <- data_matching %>%
  pivot_longer(cols = c(contains("no2"), contains("o3"), contains("pm10"), contains("pm25"), contains("so2")), names_to = "variable", values_to = "concentration") %>%
  mutate(pollutant = NA %>%
           ifelse(str_detect(variable, "no2_l"), "NO2 Longchamp",.) %>%
           ifelse(str_detect(variable, "no2_sl"), "NO2 Saint-Louis",.) %>%
           ifelse(str_detect(variable, "o3"), "O3 Longchamp",.) %>%
           ifelse(str_detect(variable, "pm10_l"), "PM10 Longchamp",.) %>%
           ifelse(str_detect(variable, "pm10_sl"), "PM10 Saint-Louis",.) %>%
           ifelse(str_detect(variable, "pm25"), "PM2.5 Longchamp",.) %>%
           ifelse(str_detect(variable, "so2"), "SO2 Lonchamp",.)) %>%
  mutate(time = 0 %>%
           ifelse(str_detect(variable, "lag_1"), -1, .) %>%
           ifelse(str_detect(variable, "lead_1"), 1, .)) %>%
  select(-variable)

# we nest the data by pollutant and time
data_reg_analysis <- data_reg_analysis %>% 
  group_by(pollutant, time) %>%
  nest()

data_reg_analysis <- data_reg_analysis %>%
  mutate(result = map(data, ~ function_reg_analysis(.))) %>%
  select(-data) %>%
  unnest(cols = c(result)) %>%
  select(-term) %>%
  rename("mean_difference" = "estimate", "ci_lower_95" = "conf.low", "ci_upper_95" = "conf.high") %>%
  mutate(analysis = "Outcome Regression")



data_neyman <- readRDS(
  here(
    "inputs",
    "1.data",
    "2.daily_data",
    "2.data_for_analysis",
    "1.cruise_experiment",
    "data_neyman.rds"
  )
)

data_neyman <- data_neyman %>%
  mutate(analysis = "Constrained Pair Matching")

data_neyman_reg <- bind_rows(data_neyman, data_reg_analysis)

# make the graph

# create an indicator to alternate shading of confidence intervals
data_neyman_reg <- data_neyman_reg %>%
  arrange(pollutant, time) %>%
  mutate(stripe = ifelse((time %% 2) == 0, "Grey", "White")) %>%
  ungroup()

# make the graph
graph_reg_analysis <-
  ggplot(
    data_neyman_reg,
    aes(
      x = as.factor(time),
      y = mean_difference,
      ymin = ci_lower_95,
      ymax = ci_upper_95,
      colour = analysis,
      shape = analysis
    )
  ) +
  geom_rect(
    aes(fill = stripe),
    xmin = as.numeric(as.factor(data_neyman_reg$time)) - 0.42,
    xmax = as.numeric(as.factor(data_neyman_reg$time)) + 0.42,
    ymin = -Inf,
    ymax = Inf,
    color = NA,
    alpha = 0.4
  ) +
  geom_hline(yintercept = 0, color = "black") +
  geom_pointrange(position = position_dodge(width = 1), size = 1.2) +
  scale_shape_manual(name = "Analysis:", values = c(16, 17)) +
  scale_color_manual(name = "Analysis:", values = c(my_orange, my_blue)) +
  facet_wrap( ~ pollutant, ncol = 4) +
  scale_fill_manual(values = c('grey90', "white")) +
  guides(fill = FALSE) +
  ylab("Average Difference \nin Concentrations (µg/m³)") + xlab("Day") +
  theme_tufte()

# display the graph
graph_reg_analysis

# save the graph
ggsave(
  graph_reg_analysis + labs(title = NULL),
  filename = here::here(
    "inputs",
    "3.outputs",
    "2.daily_analysis",
    "2.analysis_pollution",
    "1.cruise_experiment",
    "3.outcome_regression",
    "graph_reg_analysis.pdf"
  ),
  width = 30,
  height = 15,
  units = "cm",
  device = cairo_pdf
)
```













